---
title: "Simple 2D optimization"
author: "csilva"
date: "11/19/2015"
output: html_document
---

Simple test of optimization.... stocks...
At the horizon of 1 year the invariant (log-return) is distributed with V and Mu. This is the investment horizon, therefore I ignore the estimation and projection step.

```{r,echo=FALSE}
require(MASS)
Cm    = matrix(c(1,0.0,0.0,1),2,2)
Sigma = c(0.1,0.1)
V = Cm*Sigma*rep(Sigma,each=nrow(Cm)) # cov2cor
Mu= c(0,0)
R = mvrnorm(n=50000,Mu,V)
P0 = 100 # both stocks with same initial price
P  = P0*exp(R)
#P[,1] = P0
```

Is the variance matrix positive definite?

```{r, echo=FALSE}
e = eigen(V)
print(e)
```

Assume that we a buget constraint $w'1 = 1$, therefore since we have a  2 D market we can describe everything with regard to 1 weight.

```{r,echo=FALSE}
AUM_i = 1e6

w  = seq(0,1,by = 0.01)
N1 = w*AUM_i/P0  # number of shares
N2 = (1-w)*AUM_i/P0
N = cbind(N1,N2)
# Next calculate the final AUM for each state in nature
# Down the rows we have final AUM for different nature state
# The col are for different w. First col is for 1,0 .... all the way to 0,1
AUM_f = apply(N,1,function(z,P){rowSums(sweep(x = P,MARGIN = 2,STATS = z,FUN = "*"))},P=P)
```

Now we need to measure the satisfaction with different senarios. We have to choose the index of satisfaction. Let's try the utility function and certanty equivalent. Do $E[U(\phi_{\alpha})]$
```{r,echo=FALSE}
U1  = log(AUM_f)
Eu1 = colMeans(U1) # average over all the possible universes... so E[U(phi_{alpha})]
CE1 = exp(Eu1) # in units of $
gamma = AUM_i# risk aversion in $
U2 = -1*exp(-AUM_f/gamma)
Eu2 = colMeans(U2)
CE2 = -1*gamma*log(-1*Eu2)
mx1 = which.max(CE1)
plot(w,CE1,main = paste("Log utility CE vs weight of first asset",w[mx1]))
abline(v = w[mx1])
mx2 = which.max(CE2)
plot(w,CE2,main = paste("Log utility CE vs weight of first asset",w[mx2]))
abline(v = w[mx2])
```
Compare with approximation in book (5.146)

```{r,echo=FALSE}
Ephi = colMeans(AUM_f)
Vphi = apply(AUM_f,2,var)
Alog = 1/Ephi
Aexp = 1/gamma
appCe1 = Ephi-(Alog/2)*Vphi 
appCe2 = Ephi-(Aexp/2)*Vphi
mx1 = which.max(appCe1)
plot(w,appCe1,main = paste("Log utility CE vs weight of first asset",w[mx1]))
abline(v = w[mx1])
mx2 = which.max(appCe2)
plot(w,appCe2,main = paste("Log utility CE vs weight of first asset",w[mx2]))
abline(v = w[mx2])

```

How good is the log(1+x) approximation?

```{r,echo=FALSE}
x = seq(-5,5,0.01)
y = log(1+x)
z = x-0.5*x^2
plot(x,y,xlim = c(-0.5,0.5));lines(x,z,col="red")
plot(x,(y-z)/z,xlim=c(-0.5,0.5),ylim=c(-0.5,0.5))
```

Exactly like my simple computations... 1D by construction. Second asset is cash at zero IR. Therefore it is a 1 D problem

```{r,echo=FALSE}
# Log normal?
Mu1d = 0
Sig1d = 5
rg = rnorm(n = 50000,mean = Mu1d,sd = Sig1d)
s  = exp(rg)  # Indeed log normal: this is the distribution of P(t)/P(0) 
#s1 = rlnorm(n=10000,meanlog = 0,sdlog = 0.1)
L = as.matrix(s-1)  # Linear returns
wL = apply(as.matrix(w),1,function(z,L){rowSums(sweep(x = L,MARGIN = 2,STATS = z,FUN = "*"))},L=L)
U1d = log(1+wL)
EU1d = colMeans(U1d)
mx1 = which.max(EU1d)
EU1dApp = colMeans(wL)-colMeans(wL*wL)/2
EU1dApp2 = colMeans(wL)-colMeans(wL*wL)/2 + colMeans(wL*wL*wL)/3
wOpt = round(mean(L)/mean(L^2),2)
plot(w,EU1d,main=paste("BestSim=",w[mx1],"BestApp=",wOpt,"SD(wL)=",round(sd(wL),1)));lines(w,EU1dApp,col="red")
lines(w,EU1dApp2,col="blue")
abline(v = wOpt,col="red")
abline(v = w[mx1])
sig2 = Sig1d^2;
thMu = exp(Mu1d+sig2/2)-1
thSig= exp(Mu1d+sig2/2)*sqrt(exp(sig2)-1)
thS2 =(thSig^2)+((thMu+1)^2)
thL2 = thS2-2*(thMu+1)+1  # GOOD thMu/thL2 is the approx alloc
xf = seq(0,5,0.01)
f = dlnorm(xf,meanlog = 0,sdlog = 5)
plot(xf-1,f);lines(xf[-1]-1,log(xf[-1]))
```


